---
title: "sobol with Ode hw assignment"
Author: "Rosemary Juarez and Vanessa Salgado"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(deSolve)
library(sensitivity)
library(kableExtra)
```

```{r}

#===================================================================
#                           SET UP
#===================================================================


# Forest growth model
source("dforestgrowth.R")

#-------------------------------------
# Set up Sobol sensitivity analysis
# ------------------------------------

# Define parameter ranges and distributions
np <- 2000
closure_threshold <- rnorm(mean = 50, sd = 5, n = np) #parameter
K <- rnorm(mean = 250, sd = 25, n = np) #parameter
r <- rnorm(mean = 0.01, sd = 0.001, n = np) #parameter
g <- rnorm(mean = 2, sd = 0.2, n = np) #parameter

X1 <- cbind.data.frame(closure_threshold = closure_threshold, K = K, r = r, g = g)
X2 <- cbind.data.frame(closure_threshold = closure_threshold, K = K, r = r, g = g)

# Fix any negative values
X1 <- X1 %>% map_df(pmax, 0.0)
X2 <- X2 %>% map_df(pmax, 0.0)

# Create Sobol object
sens_P <- sobolSalt(model = NULL, X1, X2, nboot = 300)
colnames(sens_P$X) <- c("closure_threshold", "K", "r", "g")

# Define initial conditions and simulation times
Cinitial <- 10
simtimes <- seq(from = 1, to = 300)

#===================================================================
#                           END
#===================================================================

```


```{r}


#===================================================================
#                           CREATING FUNCTION
#===================================================================

# metric function for maximum forest size
compute_maxsize <- function(result) {
  maxsize <- max(result$C)
  return(list(maxsize = maxsize))
}

# Wrapper function:
# run ODE solver and compute maximum forest size
p_wrapper_maxsize <- function(closure_threshold, K, r, g, Cinitial, simtimes, odefunc, metricfunc) {
  parms <- list(closure_threshold = closure_threshold, K = K, r = r, g = g)
  result <- ode(y = Cinitial, times = simtimes, func = odefunc, parms = parms)
  colnames(result) <- c("time", "C")
  metrics <- metricfunc(as.data.frame(result))
  return(metrics)
}

# Run wrapper for all parameter sets
allresults_maxsize <- as.data.frame(sens_P$X) %>%
  pmap(p_wrapper_maxsize, Cinitial = Cinitial, simtimes = simtimes, odefunc = dforestgrowth, metricfunc = compute_maxsize)


#max results!
allres_maxsize <- allresults_maxsize %>% map_dfr(`[`, "maxsize")

allres_maxsize$ma

#-------------------------------
#             plot
#-------------------------------


# boxplot for maximum forest size
ggplot(allres_maxsize, aes(x = "Maximum Forest Size", y = maxsize)) +
  geom_boxplot()

# Compute Sobol indices for maximum forest size
sens_P_maxsize <- sensitivity::tell(sens_P, allres_maxsize$maxsize)

# Print Sobol indices for maximum forest size
rownames(sens_P_maxsize$S) <- c("closure_threshold", "K", "r", "g")

# first-order indices (main effect without co-variance)
sens_P_maxsize$S


# total sensitivity index
sens_P_maxsize$T

```



extra credit
```{r}
#function for forest size after 100 years
compute_size_100 <- function(result) {
  size_100 <- result$C[result$time == 100]
  return(list(size_100 = size_100))
}

# Wrapper function to run ODE solver and forest size after 100 years
p_wrapper_size_100 <- function(closure_threshold, K, r, g, Cinitial, simtimes, odefunc, metricfunc) {
  parms <- list(closure_threshold = closure_threshold, K = K, r = r, g = g)
  result <- ode(y = Cinitial, times = simtimes, func = odefunc, parms = parms)
  colnames(result) <- c("time", "C")
  metrics <- metricfunc(as.data.frame(result))
  return(metrics)
}

# Run wrapper for all parameters
allresults_size_100 <- as.data.frame(sens_P$X) %>%
  pmap(p_wrapper_size_100, Cinitial = Cinitial, simtimes = simtimes, odefunc = dforestgrowth, metricfunc = compute_size_100)

allres_size_100 <- allresults_size_100 %>% map_dfr(`[`, "size_100")

# Create boxplot for forest size after 100 years
ggplot(allres_size_100, aes(x = "Forest Size at 100 Years", y = size_100)) + geom_boxplot()

# Compute Sobol indices for forest size after 100 years
sens_P_size_100 <- sensitivity::tell(sens_P, allres_size_100$size_100)

# Print Sobol indices for forest size after 100 years
rownames(sens_P_size_100$S) <- c("closure_threshold", "K", "r", "g")
#first order
sens_P_size_100$S

#second order
sens_P_size_100$T
```

The main effect for the parameters are ...

The main effect on threshold year for the parameters are ...

# 6. Discuss meaning of results of simulation (For example think about how what parameters climate change might influence).


The maximum growth of the forest depends most on its carrying capacity, but it's also influenced by how fast it grows exponentially and then linearly. Changes in environmental factors can make the impact of carrying capacity on maximum growth even stronger. For instance, if a forest reaches its carrying capacity faster or slower due to a new climate, it will affect its growth rates and ability to reach canopy closure thresholds. The forest community can also play a role. Higher competition levels due to climate factors can impact growth rates.

The year when the forest reaches the canopy closure threshold is most affected by how fast it grows exponentially because the growth rate, denoted as 'r', is only considered when the forest is below this threshold. This sensitivity of the threshold year to the exponential growth rate suggests implications for forest growth rate because the time taken for a forest to reach canopy closure can change based on environmental conditions, which could be influenced by climate change. For example, forests experiencing more precipitation and higher temperatures would differ from those in drier conditions.

